Transfer of Learned Opponent Models in Zero Sum Games \thanks{Grants or other notes about the article that should go on the front page should be placed here. General acknowledgments should be placed at the end of the article.}

### 2021-09-30 CBB/CBB.tex
CBB manuscript

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[english,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Transfer of Learned Opponent Models in Zero Sum Games},
            pdfauthor={Ismail Guennouni1 \& Maarten Speekenbrink1},
            colorlinks=true,
            linkcolor=blue,
            filecolor=Maroon,
            citecolor=Blue,
            urlcolor=Blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{Upgrade Report}
\usepackage{csquotes}
\shorttitle{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi

\title{Transfer of Learned Opponent Models in Zero Sum Games}
\author{Ismail Guennouni\textsuperscript{1} \& Maarten Speekenbrink\textsuperscript{1}}
\date{}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Experimental Psychology, University College London}

\abstract{
Human learning transfer abilities take advantage of important cognitive building blocks such as an abstract representation of concepts underlying tasks and
causal models of the environment. One way to build abstract representations of the environment when the task involves interactions with others is to build
a model of the opponent that may inform what actions they are likely to take next. In this study, we explore opponent modelling and its transfer with the use of
computer agents with human-like limited degrees of iterated reasoning. In two experiments, we find that participants
deviate from Nash equilibrium play and learn to adapt to their opponent's strategy to exploit it. Moreover, we show that participants transfer their learning to new games and that
this transfer is moderated by the level of sophistication of the opponent, as well as the similarity between games. Computational modelling shows that players start each game with a model-based
learning strategy that facilitates between-game transfer of their opponent's strategy, but then switch to behaviour that is consistent with a model-free learning strategy in the latter stages of the interaction.
}

\begin{document}
\maketitle

@@

Being able to transfer previously acquired knowledge to a new domain is one of the hallmarks of human intelligence. This ability relies on important cognitive building blocks, such as an abstract representation of concepts underlying tasks (Lake, Ullman, Tenenbaum, \& Gershman, 2017). One way to form these representations when the task involves interactions with others, is to build a model of the person we are interacting with that offers predictions of the actions they are likely to take next. There is evidence that people learn such models of their opponents when playing repeated economic games (Stahl \& Wilson, 1995).

A model of the opponent can help increase performance in a particular game, but learning more general characteristics of an opponent may also help increase performance in other games. In this paper, we are specifically interested in the latter: How do people build and use models of their opponent to facilitate learning transfer? Repeated games, in which players interact repeatedly with the same opponent and have the ability to learn about their opponent's strategies and preferences (Mertens, 1990) are particularly useful to address this question. The early literature on learning transfer in repeated games has mostly focused on the proportion of people who play normatively optimal (e.g.~Nash equilibrium play) or use salient actions (e.g.~risk dominance) in later games, having had experience with a similar game environment previously (Camerer \& Knez, 2000; Ho, Camerer, \& Weigelt, 1998). As is well-known, a Nash equilibrium means that all players in a game act such that no-one can unilaterally improve their performance by deviating from their strategy. When playing against an opponent with a Nash-optimal strategy, I can do no better than play according to the Nash-equilibrium strategy as well. However, when faced with a player who deviates from the Nash-optimal strategy, I may be able to exploit this and also deviate from this strategy, increasing my performance beyond what is expected at Nash equilibrium. Of course, this comes with some risk, as my own deviation from Nash-optimal play may leave me open to similar exploitation.

Studies that focused on whether people can learn to exploit deviations from Nash-equilibrium play have mostly looked at the ability of players to detect and exploit action-based learning rules (Shachat \& Swarthout, 2004; Spiliopoulos, 2013). These studies used computer opponents that don't adapt play to their human opponent, mostly consisting of playing each action with a fixed probability (a mixed strategy). Findings showed that humans are capable of adapting to non-Nash-equilibrium play and detect patterns in an opponent's history of play. However, the use of mixed strategies may have limited people's ability to form accurate opponent models.
Instead of thinking of their opponents as drawing random actions from a (non-uniform) probability distribution, people may rather think of their opponents as applying a form of iterated reasoning to determine their next action. The type of reasoning we refer to takes the form of \enquote{I believe that you believe that I believe \ldots{}} and has been shown to underlie people's decisions in economic games (Camerer, 2003; Costa-Gomes, Crawford, \& Broseta, 2001). It also underlies successful models in behavioural economics, such as Level-\(k\) and Cognitive Hierarchy models (Camerer, Ho, \& Chong, 2004), which posit that people assume their opponent applies a limited level of iterative reasoning, taking actions that are best responses to those of their (modelled) opponent.

@@

Iterative reasoning strategies can explain non-equilibrium play in a range of games, such as the \(p\)-beauty contest game (Nagel, 1995), dominance solvable games (Camerer et al., 2004), and various standard normal form games (Costa-Gomes et al., 2001). In the present study, we endow our computer opponents with limited ability for iterative reasoning and assess whether (1) human players adapt their strategy to exploit this limited reasoning of their opponent, and (2) whether they are able to generalize a learned opponent model to other games. In two experiments, participants face the same opponent (Experiment 1) or the same two opponents (Experiment 2) in three consecutive games: the well-known Rock-Paper-Scissors game, the structurally similar Fire-Water-Grass game, and a less similar Numbers (Experiment 1) or Shootout (Experiment 2) game. To foreshadow our results, we find evidence that participants transfer the learned strategy of their opponent to other games, and that this transfer is moderated by the sophistication of their opponent and the similarity between the games. Moreover, using computational modelling, we find evidence that participants switch from relying on an opponent model in the early stages of the games, to a more habitual and cognitively less demanding strategy in the later stages of the games.

@@

The Fire-Water-Grass (FWG) game is identical to RPS in all but action labels: Fire burns grass, water extinguishes fire, and grass absorbs water. We use this game as we are interested in whether learning is transferred in a fundamentally similar game where the only difference is in the label of the possible actions. This should make it relatively easy to generalize knowledge of the opponent's strategy, provided this knowledge is on a sufficiently abstract level, such as knowing the opponent is a level-1 or level-2 player. Crucially, learning simple contingencies such as \enquote{If I played rock on the previous round, playing scissors next will likely result in a win}, is not generalizable to this similar game, as these contingencies are tied to the labels of the actions.

@@

The computer opponent was programmed to use either a level-1 or level-2 strategy in all the games. A level-1 player is defined as a player who best responds to a level-0 player. A level-0 player plays in a non-strategic way and does not consider their opponent's actions. Here, we assume a level-0 player simply repeats their previous action. There are other ways to define a level-0 player. For instance, as repeating their action if it resulted in a win and choosing randomly from the remaining actions otherwise, or choosing randomly from all actions. As a best response to a uniformly random action is itself a random action, defining a level-0 player in such a way would make a level-1 opponent's strategy much harder to discern. Because we are mainly interested in generalization of knowledge of an opponent's strategy to other games, which requires good knowledge of this strategy, we opted for this more deterministic formulation of a level-0 player. The level-2 computer opponent assumes in turn that the participant is a level-1 opponent, playing according to the strategy just described. To make the computer opponent's strategy not too obvious, we introduced some randomness in actions of computer opponent, making them play randomly in 10\% of all trials. Note that at all levels, the strategies are contingent on the actions taken in the previous round. The choice of this type of strategy is consistent with evidence that humans strategically use information from last round play of their opponents in zero-sum games (Batzilis, Jaffe, Levitt, List, \& Picel, 2016; Wang, Xu, \& Zhou, 2014). Table \ref{table:1} shows an example of the computer opponent's actions in response to the previous round play.

\begin{table}[bth!]
\caption{Example of how a level-1 and level-2 computer agent plays in response to actions taken in the previous round.}
\centering
\label{table:1}
\begin{tabular}{||c c | c c||} 
 \hline
 Human action $t-1$ & Computer action $t-1$  & Computer level-1 action $t$ & Computer level-2 action $t$ \\ [0.5ex] 
 \hline\hline
 Paper & Rock & Scissors & Scissors \\ 
 Scissors  & Scissors & Rock & Paper \\
 Rock & Paper & Paper & Rock \\
 ... & ... & ... & ... \\ [1ex] 
 \hline
\end{tabular}
\end{table}

@@

Participants were informed they would play three different games against the same computer opponent. Participants were told that the opponent cannot cheat and will choose its actions simultaneously with them, without prior knowledge of the participant's choice. After providing informed consent and reading the instructions, participants answered a number of comprehension questions. They then played the three games against their opponent in the order RPS, FGW, and Numbers. An example of the interface for the RPS game is provided in Figure~\ref{fig:feedback-rps-exp2}. On each round, the human player chooses an action, and after a random delay (between 0.5 and 3 seconds) is shown the action chosen by the computer opponent and the outcome of that round.A total of 50 rounds of each game was played with the player's score displayed at the end of each game. The score was calculated as the number of wins minus the number of losses. Ties did not affect the score. In order to incentivise the participants to maximise the number of wins against their opponent, players were paid a bonus at the end of the experiment proportional to their final score (each point was worth £0.02). After playing all the games, participants were asked questions about their beliefs about the computer opponent, related to whether they thought they learned their opponent's strategy, and how difficult they found playing against their opponent. They were then debriefed and thanked for their participation.

@@

{\centering \includegraphics{images/feedback_rps}

@@

\hypertarget{results}{%
\subsubsection{Results}\label{results}}

@@

{\centering \includegraphics{paper_draft_2021_files/figure-latex/exp1-avg-scores-game-1}

@@

@@

To assess learning within and between games, we estimate a linear mixed-effects model with score as the dependent variable, and Condition (Level 1 or Level 2), Game (RPS, FWG, Numbers) and Block (first-half , second-half) as independent variables, including all main effects and interactions. The model included random intercepts for participants. The results show a main effect of Game (\(F(2,250) = 12.90\), \(\eta^{2} = 0.09\), \(p < .001\)) indicating that performance differs between the games. Post-hoc pairwise comparisons show that performance in the FWG game was significantly higher than in the RPS game (\(t(100) =3.78\), \(p < .001\)) and the Numbers game (\(t(100) = 3.32\) , \(p = .002\)). The score in RPS was not significantly different from the score in Numbers (\(t(100) = 0.45\) , \(p = .65\)). The main effect of Block (\(F(2,250) = 15.16\) , \(\eta^{2} = 0.06\), \(p < .001\)) shows that the score in the first half of each game (\(M = 0.29\)) was significantly lower than in the second half (\(M = 0.40\)), which indicates within-game learning. The main effect of Condition (\(F(1,50) = 5.44\), \(\eta^{2} = 0.10\), \(p = .024\)) indicates that scores were higher against the level-1 player (\(M = 0.43\)) than against the level-2 player (\(M = 0.27\)). Thus, it appears that is was harder for participants to exploit the strategy of the more sophisticated level-2 opponent than the comparatively less sophisticated level-1 opponent. There was also an interaction between Game and Condition (\(F(2,250) = 4.52\) , \(\eta^{2} = 0.03\), \(p = .01\)). \textbf{Post-hoc pairwise comparisons show that, for participants facing level-1 opponents, scores on the FWG game were significantly higher than RPS (\(t(250) = 3.85\) , \(p = .002\)), and also significantly higher than Numbers (\(t(250) = 3.79\) , \(p = .002\)). For participants facing level-2 opponents however, there was no significant difference between scores on the three games.}

\hypertarget{learning-transfer}{%
\subsubsection{Learning transfer}\label{learning-transfer}}

@@

{\centering \includegraphics{paper_draft_2021_files/figure-latex/exp1-early-score-by-opp-1}

@@

\hypertarget{discussion}{%
\subsubsection{Discussion}\label{discussion}}

@@

Participants played three games: Rock-Paper-Scissors (RPS), Fire-Water-Grass (FWG), and the penalty Shootout game. The first two games were identical to the ones used in Experiment 1. In the Shootout game, participants took the role of the a football (soccer) player in a penalty situation, with the computer opponent taking the role of the goalkeeper. Players had the choice between three actions: shooting the football to the left, right, or centre of the goal. Similarly, the goalkeeper chooses between defending the left, right, or centre of the goal. If participants shoot in a different direction than where the goalkeeper defends, they win the round and the goalkeeper loses. Otherwise, the goalkeeper catches the ball and the player loses the round. There is no possibility of ties in this game. Figure~\ref{fig:screenshot-shootout} shows a snapshot of play in the Shootout game. What makes this game different to the other games is that there are now two ways to beat the opponent: if the shooter thinks their opponent is going to choose to defend \enquote{right} in the next round, they can win by either choosing to shoot \enquote{left} or \enquote{center}. A level-1 shooter who thinks that their goalkeeper opponent will repeat their last action has thus two possible best responses. A level-1 goalkeeper, however, has only a single best response (defending where their opponent aimed in the last round). A level-2 goalkeeper, who believes their opponent is a level-1 shooter, will have two best responses however. We programmed the level-2 computer player to choose randomly between these two best responses.

@@

{\centering \includegraphics{images/shootout}

@@

{\centering \includegraphics{paper_draft_2021_files/figure-latex/exp2-score-by-opp-1}

@@

\hypertarget{learning-transfer-1}{%
\subsubsection{Learning transfer}\label{learning-transfer-1}}

@@

{\centering \includegraphics{paper_draft_2021_files/figure-latex/exp2-early-score-by-opp-1}

@@

To measure transfer between games, we again focus on participants' scores in the initial 5 rounds after the first round (rounds 2-6) of each game and only during the first time they interact with an opponent in a game (see Figure \ref{fig:exp2-early-score-by-opp}). We estimated a linear mixed-effects model with the early round scores as the dependent variable, and Condition (level-1 first, level-2 first), Opponent type (level-1 or level-2) and Game (RPS, FWG, Shootout) as independent variables, including all main effects and interactions. The model also included random intercepts for participants. We found a main effect of Game (\(F(2,240) = 6.57\), \(p = .002\)). \textbf{The effect of game reflects early scores on FWG being significantly higher than on RPS (\(t(240) = 3.01\), \(p = .005\)) as well as scores on shootout being significantly higher than RPS (\(t(240) = 3.25\), \(p = .003\)). We also found an interaction effect between Game and Opponent type (\(F(2,240) = 3.34\), \(p = .037\)). Post-hoc pairwise comparisons (applying Holm correction for multiple comparisons) show that the difference between early scores in RPS and shootout was mostly driven by participants facing level-1 players ((\(t(240) = 4.12\), \(p < .001\))). No signifcant differences were found between opponents for other games, or between games when facing the same opponent.}

\textbf{MS: What do these effects indicate? Report post-hoc comparisons between games, and e.g.~differences between opponents for each game, here. The later tests against 0 are not adequate to understand the effects here.}

@@

\hypertarget{discussion-1}{%
\subsubsection{Discussion}\label{discussion-1}}

@@

\textbf{MS: why is there a superscript in \(h^k\) and \(r^k\)? }
\textbf{IG : Because they depend on strategy k}

\textbf{MS: And is the indicator variable \(I(o^k = o_{t})\)?}
\textbf{IG: Yes, corrected}

>>>>

% !TeX program = pdfLaTeX
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

%% load any required packages here

% Pandoc citation processing

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\title{Transfer of Learned Opponent Models in Zero Sum Games \thanks{Grants or other notes about the article that should go on the front page
should be placed here. General acknowledgments should be placed at the
end of the article.} }

\author{  Ismail Guennouni \and  Maarten Speekenbrink \and  }

\institute{
        Ismail Guennouni \at
     Department of Experimental Psychology, University College London, 26
 Bedford Way, London WC1H 0AP, United Kingdom \\
     \email{\href{mailto:i.guennouni.17@ucl.ac.uk}{\nolinkurl{i.guennouni.17@ucl.ac.uk}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Maarten Speekenbrink \at
     Department of Experimental Psychology, University College London, 26
 Bedford Way, London WC1H 0AP, United Kingdom \\
     \email{\href{mailto:m.speekenbrink@ucl.ac.uk}{\nolinkurl{m.speekenbrink@ucl.ac.uk}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
Human learning transfer abilities take advantage of important cognitive
building blocks such as an abstract representation of concepts
underlying tasks and causal models of the environment. One way to build
abstract representations of the environment when the task involves
interactions with others is to build a model of the opponent that may
inform what actions they are likely to take next. In this study, we
explore opponent modelling and its transfer with the use of computer
agents with human-like limited degrees of iterated reasoning. In two
experiments, we find that participants deviate from Nash equilibrium
play and learn to adapt to their opponent's strategy to exploit it.
Moreover, we show that participants transfer their learning to new games
and that this transfer is moderated by the level of sophistication of
the opponent (second experiment), as well as the similarity between
games. Computational modelling shows that players start each game with a
model-based learning strategy that facilitates between-game transfer of
their opponent's strategy, but then switch to behaviour that is
consistent with a model-free learning strategy in the latter stages of
the interaction.
\\
\keywords{
        Learning Transfer \and
        Opponent Modelling \and
        Hidden Markov Models \and
        Bayesian Cognitive Hierarchy \and
    }

\end{abstract}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

@@

Being able to transfer previously acquired knowledge to a new domain is
one of the hallmarks of human intelligence. This ability relies on
important cognitive building blocks, such as an abstract representation
of concepts underlying tasks (Lake et al. 2017). One way to form these
representations when the task involves interactions with others, is to
build a model of the person we are interacting with that offers
predictions of the actions they are likely to take next. There is
evidence that people learn such models of their opponents when playing
repeated economic games (Stahl and Wilson 1995).

A model of the opponent can help increase performance in a particular
game, but learning more general characteristics of an opponent may also
help increase performance in other games. In this paper, we are
specifically interested in the latter: How do people build and use
models of their opponent to facilitate learning transfer? Repeated
games, in which players interact repeatedly with the same opponent and
have the ability to learn about their opponent's strategies and
preferences (Mertens 1990) are particularly useful to address this
question. The early literature on learning transfer in repeated games
has mostly focused on the proportion of people who play normatively
optimal (e.g.~Nash equilibrium play) or use salient (e.g.~risk
dominance) actions in later games, having had experience with a similar
game environment previously (Ho, Camerer, and Weigelt 1998; Camerer and
Knez 2000). As is well-known, a Nash equilibrium means that all players
in a game act such that no-one can unilaterally improve their
performance by deviating from their strategy. When playing against an
opponent with a Nash-optimal strategy, I can do no better than play
according to the Nash-equilibrium strategy as well. However, when faced
with a player who deviates from the Nash-optimal strategy, I may be able
to exploit this and also deviate from this strategy, increasing my
performance beyond what is expected at Nash equilibrium. Of course, this
comes with some risk, as my own deviation from Nash-optimal play may
leave me open to similar exploitation.

Studies that focused on whether people can learn to exploit deviations
from Nash-equilibrium play have mostly looked at the ability of players
to detect and exploit action-based learning rules (Spiliopoulos 2013;
Shachat and Swarthout 2004). These studies used computer opponents that
don't adapt play to their human opponent, mostly consisting of playing
each action with a fixed probability (a mixed strategy) or using a
pre-determined sequence. Findings showed that humans are capable of
adapting to non-Nash-equilibrium play and detect patterns in an
opponent's history of play. However, the use of mixed strategies may
have limited people's ability to form accurate opponent models.

Instead of thinking of their opponents as drawing random actions from a
(non-uniform) probability distribution, people may rather think of their
opponents as applying a form of iterated reasoning to determine their
next action. The type of reasoning we refer to takes the form of ``I
believe that you believe that I believe \ldots{}'' and has been shown to
underlie people's decisions in economic games (Camerer 2003;
Costa-Gomes, Crawford, and Broseta 2001). It also underlies successful
models in behavioural economics, such as Level-\(k\) and Cognitive
Hierarchy models (Camerer, Ho, and Chong 2004), which posit that people
assume their opponent applies a limited level of iterative reasoning,
taking actions that are best responses to those of their (modelled)
opponent.

@@

Iterative reasoning strategies can explain non-equilibrium play in a
range of games, such as the \(p\)-beauty contest game (Nagel 1995),
dominance solvable games (Camerer, Ho, and Chong 2004), and various
standard normal form games (Costa-Gomes, Crawford, and Broseta 2001). In
the present study, we endow our computer opponents with a limited
ability for iterative reasoning and assess whether (1) human players
adapt their strategy to exploit this limited reasoning of their
opponent, and (2) whether they are able to generalize a learned opponent
model to other games. In two experiments, participants face the same
opponent (Experiment 1) or the same two opponents (Experiment 2) in
three consecutive games: the well-known Rock-Paper-Scissors game, the
structurally similar Fire-Water-Grass game, and a less similar Numbers
(Experiment 1) or Shootout (Experiment 2) game. To foreshadow our
results, we find evidence that participants transfer the learned
strategy of their opponent to other games, and that this transfer is
moderated by the sophistication of their opponent and the similarity
between the games. Moreover, using computational modelling, we find
evidence that participants switch from relying on an opponent model in
the early stages of the games, to a more habitual and cognitively less
demanding strategy in the later stages of the games.

@@

The Fire-Water-Grass (FWG) game is identical to RPS in all but action
labels: Fire burns grass, water extinguishes fire, and grass absorbs
water. We use this game as we are interested in whether learning is
transferred in a fundamentally similar game where the only difference is
in the label of the possible actions. This should make it relatively
easy to generalize knowledge of the opponent's strategy, provided this
knowledge is on a sufficiently abstract level, such as knowing the
opponent is a level-1 or level-2 player. Crucially, learning simple
contingencies such as ``If I played rock on the previous round, playing
scissors next will likely result in a win'', is not generalizable to
this similar game, as these contingencies are tied to the labels of the
actions.

@@

The computer opponent was programmed to use either a level-1 or level-2
strategy in all the games. A level-1 player is defined as a player who
best responds to a level-0 player. A level-0 player plays in a
non-strategic way and does not consider their opponent's actions. Here,
we assume a level-0 player simply repeats their previous action. There
are other ways to define a level-0 player. For instance, as repeating
their action if it resulted in a win and choosing randomly from the
remaining actions otherwise, or choosing randomly from all actions. As a
best response to a uniformly random action is itself a random action,
defining a level-0 player in such a way would make a level-1 opponent's
strategy much harder to discern. Because we are mainly interested in
generalization of knowledge of an opponent's strategy to other games,
which requires good knowledge of this strategy, we opted for this more
deterministic formulation of a level-0 player. The level-2 computer
opponent assumes in turn that the participant is a level-1 opponent,
playing according to the strategy just described. To make the computer
opponent's strategy not too obvious, we introduced some randomness in
actions of computer opponent, making them play randomly in 10\% of all
trials. Note that at all levels, the strategies are contingent on the
actions taken in the previous round. The choice of this type of strategy
is consistent with evidence that humans strategically use information
from last round play of their opponents in zero-sum games (Batzilis et
al. 2016; Wang, Xu, and Zhou 2014). Table \ref{tab:table-actions} shows
an example of the computer opponent's actions in response to the
previous round play.

\begin{table}

\caption{\label{tab:table-actions}Example of how a level-1 and level-2 computer agent plays in response to actions taken in the previous round}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llll}
\toprule
Human action t-1 & Computer action t-1 & Computer level-1 action t & Computer level-2 action t\\
\midrule
Paper & Rock & Scissors & Scissors\\
Scissors & Scissors & Rock & Paper\\
Rock & Paper & Paper & Rock\\
... & ... & ... & ...\\
\bottomrule
\end{tabular}}
\end{table}

@@

Participants were informed they would play three different games against
the same computer opponent. Participants were told that the opponent
cannot cheat and will choose its actions simultaneously with them,
without prior knowledge of the participant's choice. After providing
informed consent and reading the instructions, participants answered a
number of comprehension questions. They then played the three games
against their opponent in the order RPS, FGW, and Numbers. An example of
the interface for the RPS game is provided in Figure
\ref{fig:feedback-rps-exp2}. On each round, the human player chooses an
action, and after a random delay (between 0.5 and 3 seconds) is shown
the action chosen by the computer opponent and the outcome of that
round. A total of 50 rounds of each game was played with the player's
score displayed at the end of each game. The score was calculated as the
number of wins minus the number of losses. Ties did not affect the
score. In order to incentivise the participants to maximise the number
of wins against their opponent, players were paid a bonus at the end of
the experiment proportional to their final score (each point was worth
£0.02). After playing all the games, participants were asked questions
about their beliefs about the computer opponent, related to whether they
thought they learned their opponent's strategy, and how difficult they
found playing against their opponent. They were then debriefed and
thanked for their participation.

@@

{\centering \includegraphics[width=\textwidth]{../Report/images/feedback_rps}

@@

@@

{\centering \includegraphics[width=\textwidth]{CBB_files/figure-latex/exp1-avg-scores-game-1}

@@

\hypertarget{results}{%
\subsection{Results}\label{results}}

@@

To assess learning within and between games, we estimate a linear
mixed-effects model with score as the dependent variable, and Condition
(Level 1 or Level 2), Game (RPS, FWG, Numbers) and Block (first-half ,
second-half) as independent variables, including all main effects and
interactions. The model included random intercepts for participants. The
results show a main effect of Game (\(F(2,250) = 12.90\),
\(\eta^{2} = 0.09\), \(p < .001\)). Post-hoc pairwise comparisons show
that performance in the FWG game was significantly higher than in the
RPS game (\(t(100) =3.78\), \(p < .001\)) and the Numbers game
(\(t(100) = 3.32\) , \(p = .002\)). The score in RPS was not
significantly different from the score in Numbers (\(t(100) = 0.45\) ,
\(p = .65\)). The main effect of Block (\(F(2,250) = 15.16\) ,
\(\eta^{2} = 0.06\), \(p < .001\)) shows that the score in the first
half of each game (\(M = 0.29\)) was significantly lower than in the
second half (\(M = 0.40\)), which indicates within-game learning. The
main effect of Condition (\(F(1,50) = 5.44\), \(\eta^{2} = 0.10\),
\(p = .024\)) indicates that scores were higher against the level-1
player (\(M = 0.43\)) than against the level-2 player (\(M = 0.27\)).
Thus, it appears that is was harder for participants to exploit the
strategy of the more sophisticated level-2 opponent than the
comparatively less sophisticated level-1 opponent. There was also an
interaction between Game and Condition (\(F(2,250) = 4.52\) ,
\(\eta^{2} = 0.03\), \(p = .01\)). Post-hoc pairwise comparisons
(\(p\)-values adjusted using Holm method for multiple comparisons) show
that, for participants facing level-1 opponents, scores on the FWG game
were significantly higher than those for both RPS (\(t(250) = 3.85\) ,
\(p = .002\)), and Numbers (\(t(250) = 3.79\) , \(p = .002\)). For
participants facing level-2 opponents however, there was no significant
difference between scores on the three games.

\hypertarget{learning-transfer}{%
\subsection{Learning transfer}\label{learning-transfer}}

@@

{\centering \includegraphics[width=\textwidth]{CBB_files/figure-latex/exp1-early-score-by-opp-1}

@@

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

@@

Participants played three games: Rock-Paper-Scissors (RPS),
Fire-Water-Grass (FWG), and the penalty Shootout game. The first two
games were identical to the ones used in Experiment 1. In the Shootout
game, participants took the role of the a football (soccer) player in a
penalty situation, with the computer opponent taking the role of the
goalkeeper. Players had the choice between three actions: shooting the
football to the left, right, or centre of the goal. Similarly, the
goalkeeper chooses between defending the left, right, or centre of the
goal. If participants shoot in a different direction than where the
goalkeeper defends, they win the round and the goalkeeper loses.
Otherwise, the goalkeeper catches the ball and the player loses the
round. There is no possibility of ties in this game. Figure
\ref{fig:screenshot-shootout} shows a snapshot of play in the Shootout
game. What makes this game different to the other games is that there
are now two ways to beat the opponent: if the shooter thinks their
opponent is going to choose to defend ``right'' in the next round, they
can win by either choosing to shoot ``left'' or ``center''. A level-1
shooter who thinks that their goalkeeper opponent will repeat their last
action has thus two possible best responses. A level-1 goalkeeper,
however, has only a single best response (defending where their opponent
aimed in the last round). A level-2 goalkeeper, who believes their
opponent is a level-1 shooter, will have two best responses however. We
programmed the level-2 computer player to choose randomly between these
two best responses.

@@

{\centering \includegraphics[width=\textwidth]{../Report/images/shootout}

@@

{\centering \includegraphics[width=\textwidth]{CBB_files/figure-latex/exp2-score-by-opp-1}

@@

\hypertarget{learning-transfer-1}{%
\subsection{Learning transfer}\label{learning-transfer-1}}

To measure transfer between games, we again focus on participants'
scores in the initial 5 rounds after the first round (rounds 2-6) of
each game and only during the first time they interact with an opponent
in a game (see Figure \ref{fig:exp2-early-score-by-opp}). We estimated a
linear mixed-effects model with the early round scores as the dependent
variable, and Condition (level-1 first, level-2 first), Opponent type
(level-1 or level-2) and Game (RPS, FWG, Shootout) as independent
variables, including all main effects and interactions. The model also
included random intercepts for participants. We found a main effect of
Game (\(F(2,240) = 6.57\), \(p = .002\)). The effect of game reflects
early scores on FWG being significantly higher than on RPS
(\(t(240) = 3.01\), \(p = .005\)) as well as scores on shootout being
significantly higher than RPS (\(t(240) = 3.25\), \(p = .003\)). We also
found an interaction effect between Game and Opponent type
(\(F(2,240) = 3.34\), \(p = .037\)). Post-hoc pairwise comparisons
(p-values adjusted using Holm method for multiple comparisons) show that
the difference between early scores in RPS and shootout was mostly
driven by participants facing level-1 players ((\(t(240) = 4.12\),
\(p < .001\))). No signifcant differences were found between
participants facing different opponents in the same game, or between
games when facing the same opponent.

@@

{\centering \includegraphics[width=\textwidth]{CBB_files/figure-latex/exp2-early-score-by-opp-1}

@@

@@

\hypertarget{discussion-1}{%
\subsection{Discussion}\label{discussion-1}}

@@

For more details on the self-tuning EWA model, we refer the reader to
(Ho, Camerer, and Chong 2004). As in the RL model above, actions are
chosen based on a softmax decision rule:
